### 2015-02-07 common/lstlang-bomil.tex
Initial commit

\lstdefinelanguage{BOMIL}{%
  morekeywords={%
    addr, alloc, and, any, apply, bool, byte, case, ccall, cont, define, do, double,%
    end, enum, extern, false, float, fun, host_vproc, if, in, inline, int, let, long, module,%
    noreturn, promote, pure, return, short, tag, throw, true, typedef, unwrap, use_rw,%
    vec128, void, vproc, vpload, vpstore, wrap,%
  },%
  keywordsprefix=@,
  % otherkeywords={(,),[,],\{,\},\,,:,...,_,|,=,=>,->,\#,:>},
  sensitive,%
  morecomment=[n]{(*}{*)},%
  morestring=[d]",%
}[keywords,comments,strings]%