### 2015-02-07 icfp15-servo/libraries.tex

%!TEX root = paper.tex

\section{Libraries and abstractions}
\label{sec:libraries}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: