### 2015-02-16 icfp15-servo/safety.tex

%!TEX root = paper.tex

\section{Unsafe code}
\label{sec:safety}

There are several categories of code in Servo that cannot be written in the safe subset of Rust.

\subsection{FFI calls}

Interoperating with native C code often requires raw access to memory allocated in Rust or returns
pointers to memory that was allocated by the C libraries.
Translations to an from these raw pointers requires the use of unsafe code.

\subsection{Intrusive data structures}

Rust's ownership model assumes that there is a single owner for each piece of data.
However, many data structures do not follow that model, in order to provide asymptotic
improvements in access.
For example, a doubly-linked list contains a back pointer to the previous element to aid
in traversals in the opposite direction.
Many optimized hashtable implementations also have both hash-based access to items
and a linked list of all of the keys or values.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: