### 2015-02-27 icfp15-servo/abstract.tex

%!TEX root = paper.tex
%
\begin{abstract}
All modern web browsers --- Internet Explorer, Firefox, Chrome, Opera, and Webkit ---
have a core rendering engine written in \Cplusplus.
This language choice was made because it affords the systems programmer complete control
of the underlying hardware features and memory in use and it provides a transparent
compilation model.

Servo is a project started at Mozilla Research to build a new web browser engine that
preserves the capabilities of these other browser engines but also both takes more
advantage of the recent trends in parallel hardware and is more memory-safe.
We use a new language, Rust, that provides us a similar level of control of the underlying
system but which builds on many concepts familiar to the functional programming community
to form a novelty - a useful, safe systems programming language.

In this paper, we show how a language with an affine type system and many syntactic
features familiar to functional language programmers can be successfully used to build
state-of-the-art systems software.
We also outline several pitfalls encountered along the way and describe some potential
areas for future research.
\end{abstract}