### 2015-02-27 icfp15-servo/browsers.tex
Edits from Simon

%!TEX root = paper.tex

\section{Browsers}
\label{sec:browsers}

Modern web browsers do not just load static pages, but can also handle pages that have similar
complexity to native applications.
From application suites such as the Google Apps\footnote{\url{https://apps.google.com}} to games
based on the Unreal Engine\footnote{\url{https://www.unrealengine.com/}},
modern browsers have the ability to handle much more than simple static pages.
\figref{fig:browser} shows the steps in processing a site.
While the naming is specific to the Servo browser, similar steps are used in all modern browsers.\footnote{\url{www.html5rocks.com/en/tutorials/internals/howbrowserswork/}}
\begin{figure*}[ht]
  \begin{center}
    \includegraphics[scale=0.7]{pics/browser}
  \end{center}%
  \caption{Processing stages and intermediate representations in a browser engine.}
  \label{fig:browser}
\end{figure*}%

\subsection{Parsing HTML and CSS}

A URL identifies a resource to load.
This resource usually consists of HTML, which is then parsed and typically turned into a Document Object
Model (DOM) tree.
From a programming languages standpoint, there are several interesting aspects of the parser design
for HTML.
First, though the specification allows the browser to abort on a parse error\footnote{\url{https://html.spec.whatwg.org/multipage/syntax.html#parsing}},
in practice browsers follow the recovery algorithms described in that specification precisely so that
even ill-formed HTML will be handled in an interoperable way across all browsers..
Second, due to the presence of the \lstinline[language=HTML]{<script>} tag, the token stream can be modified
during operation.
For example, the below example that injects a close tag for the script block works in all modern browsers.
\begin{lstlisting}[language=HTML]
<html>
  <script>
  document.write(``<h'');
  </script>1>
  This is a h1 title

  <script>
  document.write(``<!-'');
  </script>-
  This is commented
</html>
\end{lstlisting}
This requirement causes parsing to pause while JavaScript code is running.
But, since resource loading is such a large factor in the latency of loading many webpages (particularly on mobile),
all modern parsers also perform token stream scanning and prefetch of resources likely to be required~\cite{browsers-slow-smartphones}.

\subsection{Layout}

After constructing the DOM and determining the set of styles defined in the Cascading Style Sheets (CSS) that has
been loaded, those styles are applied to the DOM and a new tree that represents the elements is created.
This process can create many more elements than previously existed in the DOM --- for example, when a list item is
styled to have an associated counter glyph.

\subsection{Display}

That tree of elements, called the flow tree in Servo, is then processed to product a set of display list items.
These list items are the actual graphical elements, text runs, etc. in their final on-screen positions.
The order in which these elements are displayed is well-defined by the standard\footnote{\url{http://www.w3.org/TR/CSS21/zindex.html#painting-order}}.

\subsection{Rendering}

Once all of the elements to appear on screen have been computed, these elements are rendered, or painted, into memory buffers or
directly to graphics surfaces.

\subsection{Scripting}

Whether through timers, \lstinline[language=HTML]{<script>} blocks in the HTML, user interactions, or other event handlers,
JavaScript code may execute at any point during parsing, layout, and painting or afterwards during display.
These scripts can modify the DOM tree, which may require rerunning the layout and painting passes in order to update
the output.
Most modern browsers use some form of dirty bit marking to attempt to minimize the recalculations during this process.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "paper"
%%% End: