### 2015-04-22 icfp15-servo/response.txt
Edit response

We would like to thank the reviewers for their time spent reviewing the paper and providing their feedback. We agree with the presentation-related comments, in particular that there is writing that could be improved and we could use a better presentation of Rust, and will incorporate this feedback in a final draft.

In the following, text preceded by a '-' is from a reviewer's question or comment.

Review 1

- Section 4.2, what does it means that Rust embeds serialized generic function
ASTs within "compiled" binary? What's the point? When and how these ASTs
are later used?

Rust implements generic functions by generating optimized machine code for each
type instantiation. When a library exports a generic function, the compiler
will include a serialized AST in the "compiled" output to support instantiation
by library consumers. We will clarify this in the paper.

Similarly, C++ libraries often contain monomorphic code that is compiled once,
in addition to templates that are exposed in header files. Rust improves
usability by performing this split automatically.

- Section 5.1, JS JIT, I'm not sure to understand why you consider than JIT
correctness is a problem faced by Rust, inside the Servo implementation.
A little bit of elaboration would have helped me to understand the
problem you want to address here.

We agree that JIT correctness is a problem regardless of implementation language. Indeed, we mention it due to the history of security holes resulting from mis-optimizations in SpiderMonkey. The open problem for Servo is how to obtain Rust-like memory-safety guarantees on the JIT compiler's output. We will clarify this in the final draft.

Review 2

In a final draft, we would be happy to provide a more detailed breakdown of Servo's usage of unsafe Rust code, categorized by purpose.

It's harder to quantify the cost/benefit tradeoff of Rust's static checking. Gecko is deployed on hundreds of millions of diverse end-user systems, with the result that Mozilla engineers spend substantial time tracking down individually-rare memory-safety bugs. For this reason, we believe Rust's up-front checking will be especially valuable for a browser engine.

We also did not go into great depth on our usage of Rust's memory management abstractions because we largely use our own. Because Rust does not support integer type specialization (https://github.com/rust-lang/rfcs/pull/884), many of the web-specific optimizations made to optimize default sizes and layouts of data structures based on common web usage patterns require explicit encodings.