### 2015-10-11 icse16-servo/paper.tex
Beginning the paper for ICSE16

\documentclass[preprint]{../common/acm_proc_article-sp}

\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{url}
\usepackage{graphicx}

\usepackage{tikz}
\usepackage{tkz-graph}
\usetikzlibrary{arrows}
\usetikzlibrary{fit}					% fitting shapes to coordinates
\usetikzlibrary{backgrounds}	% drawing the background after the foreground
\usetikzlibrary{arrows}

\usepackage{color}
\definecolor{Red}{rgb}{0.9,0.0,0.0}  % fixme
\definecolor{Green}{rgb}{0.0,0.4,0.0}
\definecolor{Blue}{rgb}{0.0,0.0,0.9}
\definecolor{DarkBlue}{rgb}{0.0,0.0,0.75}
\definecolor{Midnight}{rgb}{0.0,0.0,0.5}
\definecolor{Purple}{rgb}{0.5,0.0,0.4}
\definecolor{Black}{rgb}{0.0,0.0,0.0}
\definecolor{Yellow}{rgb}{1.0,1.0, 0.25}
\definecolor{Cyan}{rgb}{0.25,1.0, 1.0}

\newcommand{\cdColor}{Black}
\newcommand{\kwColor}{DarkBlue}
\newcommand{\comColor}{Red}

\usepackage{listings}
\usepackage{textcomp}
\input{../common/lstlang-rust}
\lstset{
  basicstyle=\ttfamily\color{\cdColor},
  keywordstyle=\color{\kwColor}\bfseries,
  commentstyle=\color{\comColor}\itshape,
  language=Rust,
  upquote=true
}

\newcommand{\dcfa}{{$\Delta$}{CFA}}
\newcommand{\gcfa}{{$\Gamma$}{CFA}}

\usepackage{algorithm}
\usepackage{algpseudocode}

\input{../common/common-defs}
\newcommand{\superbeta}{Super-\ensuremath{\beta{}} }

\usepackage{../common/code}

\title{Engineering the Servo Web Browser Engine using Rust}

\numberofauthors{6}

\author{
\alignauthor
Brian Anderson\\
       \affaddr{Mozilla Research}\\
       \email{banderson@mozilla.com}
\alignauthor
Lars Bergstrom\\
       \affaddr{Mozilla Research}\\
       \email{larsberg@mozilla.com}
\alignauthor
Manish Goregaokar\\
       \affaddr{Indian Institute of Technology Bombay}\\
       \email{manishg@iitb.ac.in}
\and
\alignauthor
Josh Matthews\\
       \affaddr{Mozilla}\\
       \email{jdm@mozilla.com}
\alignauthor
Keegan McAllister\titlenote{Work performed while employed at Mozilla}\\
       \email{mcallister.keegan@gmail.com}
\alignauthor
Jack Moffitt\\
       \affaddr{Mozilla}\\
       \email{jack@mozilla.com}
\and
\alignauthor
Simon Sapin\\
       \affaddr{Mozilla}\\
       \email{ssapin@mozilla.com}
}

\date{\today}

\begin{document}

\maketitle

%\thispagestyle{empty}

%\sloppy

\input{abstract}

% A category with the (minimum) three required fields
\category{H.5.4}{Information Interfaces and Presentation}{Hypertext/Hypermedia}
%A category including the fourth, optional field follows...
\category{D.2.11}{Software Engineering}{Software Architectures}[languages, patterns]

\keywords{browser engine, Rust, Servo, concurrency, parallelism} % NOT required for Proceedings

\input{intro}
\input{browsers}
\input{rust}
\input{servo}
\input{open}
\input{rel}

\section{Acknowledgments}
The Servo project was started at Mozilla Research, but has benefitted from
contributions from Samsung, Igalia, and hundreds of volunteer community members.
These contributions amount to more than half of the commits to the project,
and we are grateful for both our partners' and volunteers' efforts to make
this project a success.

We would like to thank Robert Clipsham, David Herman, Tim Kuehn, Lindsey Kuper,
Ms2ger, and Ben Striegel for their comments on drafts of this paper.

\bibliographystyle{../common/alpha}
\bibliography{../common/strings-short,../common/manticore,../common/browser}

\end{document}