### 2015-10-23 icse16-servo/abstract.tex
Additional ICFP comment

%!TEX root = paper.tex
%
\begin{abstract}
All modern web browsers --- Internet Explorer, Firefox, Chrome, Opera, and Safari ---
have a core rendering engine written in \Cplusplus{}.
This language choice was made because it affords the systems programmer complete control
of the underlying hardware features and memory in use, and it provides a transparent
compilation model.
Unfortunately, this language is both complex and highly susceptible to memory safety
issues that potentially lead to security holes.

Servo is a project started at Mozilla Research to build a new web browser engine that
preserves the capabilities of these other browser engines but also both takes 
advantage of the recent trends in parallel hardware and is more memory-safe.
We use a new language, Rust, that provides us a similar level of control of the underlying
system to \Cplusplus{} but which builds on many concepts familiar to the functional programming
community.

In this paper, we show how a language with an advanced type system can address many of
the most common security issues and software engineering challenges in other browser
engines, while still producing code that has the same performance and memory profile
required by these applications.
We also outline several pitfalls encountered along the way and describe some potential
areas for future research.
\end{abstract}