### 2015-10-24 icse16-servo/concl.tex
Add conclusion. Fix typos.

\section{Conclusion}
\label{sec:concl}
In this experience report, we have described our experiences using the new Rust
programming language to develop Servo.
Based on our experiences to date, we would recommend that either any new systems
project or new submodule within a larger C or \Cplusplus{} systems project strongly
consider the use of Rust as an implementation language.