### 2016-02-08 icse16-servo/concl.tex

\section{Conclusion}
\label{sec:concl}
In this experience report, we have described our experiences using the new Rust
programming language to develop Servo.
We cannot yet quantify the amount of additional work that is required by
the stricter typesystem to author Rust code instead of C or \Cplusplus{}.
But, we have found that the Rust language is accessible to new developers who
only have experience in \Cplusplus{}, Python, or JavaScript --- even for writing
core performance- and memory-sensitive libraries.
Further, with a relatively small team and short time, we have written a browser
engine that is increasingly competitive both in performance and functionality
with other modern browser engines.

The strongest evidence that Rust is not too expensive to write, though, is that
it can statically verify the absence of memory safety errors.
Other systems projects --- including all currently shipping browser engines --
have spent great amounts of time writing and reviewing their C or \Cplusplus{}
code, written fuzzers, and used internal and external static analysis tools, but
their security vulnerabilities are still primarily memory safety errors.
For projects where memory safety errors need to be eliminated, Rust is a
superior choice.